---
title: "HCM images - CINE"
author: "Sevde Etoz Niemeier"
date: "12/11/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
library(readxl)
library(writexl)
all_image <- read_xls("/Users/sevdeetoz/OneDrive - Johns Hopkins/current projects/HCM/ImageTable-3.xls" )

```

## Import and clean the data
Metadata information per image - not slice. Table name: "ImageTable-3.xls"
Variables named InversionTime and cardiac num of cycles should be numeric, rather than character. THis will help with the conditions like InversionTime ~ 300 (this is a typical LGE inversion time), so we can exclude inversion time of 1000 or something like that. The same for cardiac num of frames: we need cardiac num of frames > 1 for cine. It is hard to do all of this with a character array. When you convert a character to numeric in R, "Missing" will be NA and the numeric values will stay the same
```{r all_image}
#summary(all_image)
Patients <- unique(all_image$`Patient Number`)
all_image$InversionTime <- as.numeric(all_image$InversionTime)
all_image$CardiacNumofFrames <- as.numeric(all_image$CardiacNumofFrames)

```
## Get potential cine images with the first criterion. 
First criterion is: cine contrast  = missing & IR = missing & cardiac num of frames > 1
CardiacNumofFrames is a character array, so I needed to test all possibilities. WHen I only did > 0 , it also took Missing as >0. Therefore, I am trying to eliminate all options that are NOT "cardiac num of frames > 1"
Contrast data is very inhomogeneous (see below), I will also add "No" as a measure of lacking contrast. There are others I'm also not completely sure, like "omniscan" or "Default". 
```{r cine, echo=TRUE}
#get all potential cine
#For cine contrast  = missing & IR = missing & cardiac num of frames > 1

cine1 <- all_image[((all_image$Contrast == "Missing")|(all_image$Contrast == "No"))&(is.na(all_image$InversionTime))&((!is.na(all_image$CardiacNumofFrames))&(all_image$CardiacNumofFrames > 1)), ]

unique(all_image$Contrast)
```
### Cine Plots for the first criterion
EchoTime, RepetationTime, FlipAngle per image for cine images

```{r cine_plot, echo=TRUE}
plot(cine1$EchoTime)
plot(cine1$RepetitionTime)
plot(cine1$FlipAngle)
```
### Other metadata for cine with first criterion
SequenceName, SequenceVariant, ScanningSequence for cine. 
```{r cine2, echo = TRUE}
unique(cine1$SequenceVariant)
unique(cine1$SequenceName)
unique(cine1$ScanningSequence)
```
## Second criterion
The first criterion did not narrowed down to the cine (see RepetitionTime, EchoTime plots, and SequenceName/Variant etc.) There is still variation in all plots and datasets.
Cine is supposed to be a Gradient Echo, see http://www.mri-q.com/beating-heart-movies.html
Therefore, I will use ScanningSequence tag to get only the cine images. New criterion will be [first criterion & ScanningSequence == "GR"]
GR: gradient recalled. 

```{r second_criterion_cine }
cine2 <- cine1[(cine1$ScanningSequence == "GR"), ]
```
### Cine Plots for the second criterion
EchoTime, RepetationTime, FlipAngle per image for cine images
```{r second_criterion_cine_plot}
plot(cine2$EchoTime)
plot(cine2$RepetitionTime)
plot(cine2$FlipAngle)
unique(cine2$SequenceVariant)
unique(cine2$SequenceName)
unique(cine2$ScanningSequence)
```
## Third criterion: Use SequenceName
According the the document that Edem sent us (https://livejohnshopkins-my.sharepoint.com/:b:/r/personal/setozni1_jh_edu/Documents/Shared-HCM%20project/papers/BH%20ADULT%20Z4.pdf?csf=1&web=1&e=mqskMD)
cine sequences are always Trufi (which is a gradient sequence) and they start with "tfi" letters (in sequence name). In the current cine dataset, there are sequence names that start with "fl.....". I think these are flow images (phase contrast), not cine. Search for fl_ in the Edem's PDF. They also have cardiac number of images > 1. They look like cine. The flow images will also have a second image with identical metadata (except mag/phase tag) because phase contrast sequence exports two images (magnitude and phase). I look at the sequence name and the second noisy image to see them in the dicom viewer.
Here I will add the SquenceName tag in the third criterion. 
So third criterion will be = second criterion & SequenceName = "tfi..."
```{r third _criterion}
cine3 <- cine2[(substr(cine2$SequenceName,1,4) == "*tfi")|(substr(cine2$SequenceName,1,3) == "tfi"), ]
```
### Cine Plots for the third criterion
```{r third_criterion_cine_plot}
plot(cine3$EchoTime)
plot(cine3$RepetitionTime)
plot(cine3$FlipAngle)
unique(cine3$SequenceVariant)
unique(cine3$SequenceName)
unique(cine3$ScanningSequence)

```
## Final Cine Images
We need to pick some images randomly and look if they are really cine. 
```{r export_cine}
write_xlsx(cine3,"/Users/sevdeetoz/OneDrive - Johns Hopkins/current projects/HCM/finalCINE.xlsx")

```
## Get potential LGE images with the first criterion.
The sequence document that Edem sent us includes multiple LGE sequence types (check sequences that has "MDE")
SA MDE TRUFI STACK: 2D, TR~700, TE~1, InversionTime~300, SeqName: tfi, cine = off
4CH MDE TRUFI: Similar to the previous one. 
2CH MDE TRUFI: Similar to first two. 
SA MDE TFL: 2D, TR~700, TE~3, InversionTime~300, SeqName: fl, cine = off
3CH MDE TFL: Similar to previous one. 

Some LGE seems like separate 2D images from a volume, while some are just 3D volume all together. In both cardiac Number of Images=1. 

First criterion: 
(contrast != "Missing" OR contrast = "Yes") AND (500 >InversionTime > 100 AND InversionTime != "Missing") (CardiacNumofFrames = 1)

Here, inversion time of 100 and 500 is somewhat arbitrary. Ideally Inversion time should be~300, but that's not the case all the time. We will see the variation with graphs. 
```{r LGE1, echo=TRUE}
LGE1 <- all_image[((all_image$Contrast != "Missing")|(all_image$Contrast == "Yes"))&(!is.na(all_image$InversionTime)&(all_image$InversionTime > 100)&(all_image$InversionTime < 500))&(all_image$CardiacNumofFrames == 1), ]

```
### LGE Plots for the first criterion
```{r LGE1_plot}
plot(LGE1$EchoTime)
plot(LGE1$RepetitionTime)
plot(LGE1$FlipAngle)
unique(LGE1$SequenceVariant)
unique(LGE1$SequenceName)
unique(LGE1$ScanningSequence)
```
## Second criterion for LGE
I am not sure, if this is all LGE. TRs seems like they are clustered at different values. Since LGE is supposed to be a gradient echo sequence, I will remove the images with "SE" (spin echo) in the Scanning Sequence tag. 
LGE2 = LGE1& ScanningSeq != SE & IR\\SE
```{r LGE2, echo=TRUE}
LGE2 <- LGE1[(LGE1$ScanningSequence != "SE") & (LGE1$ScanningSequence != "IR\\SE") , ]
```
### LGE Plots for the second criterion
```{r LGE2_plot}
plot(LGE2$EchoTime)
plot(LGE2$RepetitionTime)
plot(LGE2$FlipAngle)
unique(LGE2$SequenceVariant)
unique(LGE2$SequenceName)
unique(LGE2$ScanningSequence)
```
LGE2 did not change the results. I am not sure why the data seem inhomogeneous. It might be because of having both 3D and 2D LGE images. Apparently, to find the most accurate inversion time for nulling the myocardium, they run multiple 2D sequences with varying inversion times. 
"In practice, TI varies with time after Gd injection, and so operators may need to adjust TI several times during the acquisition of LGE-MRI. This will prolong the scan time. Recently, a new LGE sequence called phase-sensitive IR (PSIR) has been developed to address this problem. The PSIR sequence is designed in a way that it nulls the normal myocardium consistently over a range of TIs. Operators only need to use a default TI for the whole acquisition of LGE-MRI without adjusting it." from (doi: 10.6515/ACS20150616A)
So we have a very inhomogenous LGE dataset. There are 2D, 3D LGEs, also we have LGE images with suboptimal inversion times (while they were trying to find the optimum inversion to null myocardium). Also see the inversion times above. The LGEs with suboptimal inversion times will have different contrast than "good LGEs", it might be a good idea to exclude them or adjust hem in a way to match the contrast between scar and normal myocardium. In a good LGE, myocardium should look black. I am not sure how we can do this with just the metdata information from the image table. Some ideas: look at the min and max intensity values in metadata. Or we can talk with Hayley and Dan, they might have had a similar issue. 

I am exporting the current LGEs, use with caution. 
```{r export_LGE, echo=TRUE}
write_xlsx(LGE2,"/Users/sevdeetoz/OneDrive - Johns Hopkins/current projects/HCM/finalLGE.xlsx")
```
