---
title: "HCM images"
author: "Sevde Etoz Niemeier"
date: "12/9/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
library(readxl)
library(writexl)
all_image <- read_xls("ImageTable-3.xls" )

```

## Data Summary and first criterion
Metadata information per image - not slice. Table name: "ImageTable-3.xls"
```{r all_image}
#summary(all_image)
Patients <- unique(all_image$`Patient Number`)

```
## Get potential cine images with the first criterion. 
First criterion is: cine contrast  = missing & IR = missing & cardiac cycle > 1
CardiacNumofFrames is a character array, so I needed to test all possibilities. WHen I only did > 0 , it also took Missing as >0. 
```{r cine}
#get all potential cine
#For cine contrast  = missing & IR = missing & cardiac num of frames > 1

cine1 <- all_image[(all_image$Contrast == "Missing")&(all_image$InversionTime == "Missing")&(all_image$CardiacNumofFrames!= "Missing")&(all_image$CardiacNumofFrames!= "0")&(all_image$CardiacNumofFrames> "1"), ]

```
### Cine Plots for the first criterion
EchoTime, RepetationTime, FlipAngle per image for cine images

```{r cine_plot, echo=TRUE}
plot(cine1$EchoTime)
plot(cine1$RepetitionTime)
plot(cine1$FlipAngle)
```
### Other metadata for cine with first criterion
SequenceName, SequenceVariant, ScanningSequence for cine. 
```{r cine2, echo = TRUE}
unique(cine1$SequenceVariant)
unique(cine1$SequenceName)
unique(cine1$ScanningSequence)
```
## Second criterion
The first criterion did not narrowed down to the cine (see RepetitionTime, EchoTime plots, and SequenceName/Variant etc.) There is still variation in all plots and datasets.
Cine is supposed to be a Gradient Echo, see http://www.mri-q.com/beating-heart-movies.html
Therefore, I will use ScanningSequence tag to get only the cine images. New criterion will be [first criterion & ScanningSequence == "GR"]
GR: gradient recalled. 

```{r second_criterion_cine }
cine2 <- cine1[(cine1$ScanningSequence == "GR"), ]
```
### Cine Plots for the second criterion
EchoTime, RepetationTime, FlipAngle per image for cine images
```{r second_criterion_cine_plot}
plot(cine2$EchoTime)
plot(cine2$RepetitionTime)
plot(cine2$FlipAngle)
unique(cine2$SequenceVariant)
unique(cine2$SequenceName)
unique(cine2$ScanningSequence)
```
## Third criterion: Use SequenceName
According the the document that Edem sent us (https://livejohnshopkins-my.sharepoint.com/:b:/r/personal/setozni1_jh_edu/Documents/Shared-HCM%20project/papers/BH%20ADULT%20Z4.pdf?csf=1&web=1&e=mqskMD)
cine sequences are always Trufi (which is a gradient sequence) and they start with "tfi" letters (in sequence name). In the current cine dataset, there are sequence names that start with "fl.....". I think these are flow images (phase contrast), not cine. Search for fl_ in the Edem's PDF. They also have cardiac number of images > 1. They look like cine. The flow images will also have a second image with identical metadata (except mag/phase tag) because phase contrast sequence exports two images (magnitude and phase). I look at the sequence name and the second noisy image to see them in the dicom viewer.
Here I will add the SquenceName tag in the third criterion. 
So third criterion will be = second criterion & SequenceName = "tfi..."
```{r third _criterion}
cine3 <- cine2[(substr(cine2$SequenceName,1,4) == "*tfi"), ]
```
### Cine Plots for the third criterion
```{r third_criterion_cine_plot}
plot(cine3$EchoTime)
plot(cine3$RepetitionTime)
plot(cine3$FlipAngle)
unique(cine3$SequenceVariant)
unique(cine3$SequenceName)
unique(cine3$ScanningSequence)

```
## Final Cine Images
We need to pick some images randomly and look if they are really cine. 
```{r}
write_xlsx(cine3,"/Users/sevdeetoz/OneDrive - Johns Hopkins/current projects/HCM/finalCINE.xlsx")

```
